---
title: "p8105_hw6_kw3180"
author: "Kino Watanabe"
date: "2025-11-28"
output: github_document
---

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 2

### Load the data

```{r}
library(modelr)
library(p8105.datasets)
data("weather_df") 

weather_df <- weather_df |> 
  filter(name == "CentralPark_NY")
```
What do the distribution of variables prcp and tmin look like?

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point() 

weather_df |> 
  ggplot(aes(x = prcp, y = tmax)) + 
  geom_point() 
```

### Let's do bootstrapping

```{r}
set.seed(1)

bootstrap_results = 
  weather_df |> 
  bootstrap(n = 5000) |> 
  mutate(
    df = map(strap, as_tibble),
    fits = map(df, \(df) lm(tmax ~ tmin + prcp, data = df)),
    coefs = map(fits, broom::tidy),
    rsquare = map(fits, broom::glance)
  ) |> 
  select(.id, coefs, rsquare) 


coefs_tibble <- bootstrap_results |> 
  unnest(coefs) |> 
  select(.id, term, estimate) |> 
  pivot_wider(
    names_from = term, 
    values_from = estimate) |> 
  mutate(beta_slope = tmin / prcp)


rsquare_tibble <- bootstrap_results |> 
  unnest(rsquare) |> 
  select(.id, r.squared, adj.r.squared)

bootstrap_results_merged <- 
  left_join(coefs_tibble, rsquare_tibble, by = ".id")


bootstrap_results_merged
```

### Plot the distribution after bootstrapping

```{r}
# r² distribution
rsquare_plot <- bootstrap_results_merged |> 
  ggplot(aes(x = r.squared)) +
  geom_density() +
  labs(
    x = "R-square hat estimates",
    y = "Density",
    title = "Bootstrap Distribution of r² hat")

rsquare_plot

# β¹ ÷ β² distribution
slope_plot <- bootstrap_results_merged |> 
  ggplot(aes(x = beta_slope)) +
  geom_density() +
  labs(
    x = "β1 hat/ β2 hat estimates",
    y = "Density",
    title = "Bootstrap Distribution of β1 hat/ β2 hat")

slope_plot
```

* The r-square is generally normally distributed after bootstrapping, but the distribution of r-squares are still slightly left skewed.
* The 95% CI for the proportion of variance in Central Park's tmax that is explained by predictors tmin and prcp is between 0.89 and 0.93.

* The distribution after bootstrapping for β1 hat/ β2 hat is extremely skewed to the left, and the wide distribution is reflected in its large 95% CI = (-5616.45, 4586.92). his occurs because the beta coefficient for prcp is close to 0, while the beta coefficient for tmin is close to 1. Small differences in beta 2 (prcp) lead to large changes in the slope ratio, making the estimate unstable. Additionally, precipitation shows outliers and does not appear to have a meaningful linear association with tmax, further contributing to variability in its estimated coefficient and instability in the ratio.


### 95% CI
```{r}
bootstrap_results_merged |> 
  summarize(
    r2_ci_lower = quantile(r.squared, 0.025),
    r2_ci_upper = quantile(r.squared, 0.975),
    beta_ci_lower = quantile(beta_slope, 0.025),
    beta_ci_upper = quantile(beta_slope, 0.975)
  )
```

